---
title: "InVEST Habitat Quality"
author: "Sarah Halperin"
date: "5/11/2020"
output: pdf_document
---

## Habitat Quality inVEST Model Data Collection Notes and Model Running
5-11-2020

Following Gibson (Quintas-Soriano et al., 2020) selected threats to be agriculture, low/med/high urban, primary roads, and secondary roads, and trails. 

Weights and effective distance taken from Polasky et al. 2011 appendix information. This include Ag, urban, primary, secondary, and light roads. I equated light roads with trails. 
Weights and effective distance for medium and low density taken from Blumstein and Thompson 2015. 

Sensitivity parameters were primarily taken from Blumstein and Thompson who used sensitivity for genral terrestrial species. However, these did not include primary, secondary, and trails. I have taken those from Polasky et al., 2011. 


In cases where there was distinction based on ownership I took public ownership numbers (Polaskey et al., 2011). 

To determine where habitat or not, I used Blumstein which use a binary of either habitat or not for terrestrial species and onyl had a reduction for agriculture. Therefore all non-habitat was assigned the maximimum degradation score on the landscape. With this technique need to invert the results so that 0 represents most degraded. It is possible to run for different types of species as Polaskey as numbers already and then average. 

#-------------------------------------------------------------#
#Threat Shapefiles 
Agriculture + Urban: Created raster for presence/absence from NLCD


Roads 
Using U.S. Census data you can download all roads for an individual county and then I subset the data into primary and secondary roads. This isn't that many roads should we include city roads etc.? Data:https://www.census.gov/cgi-bin/geo/shapefiles/index.php?year=2016&layergroup=Roads 

See Roads.R for creation of data inputs. 

Used data for 2016

Trails Data
Same as used in Recreation Model. Idaho.gov and urban trails from ridge to rivers. 

Data: /Volumes/Research/JodiBrandt/Common/SarahHalperin/Data/HabitatQuality_Data/Threats
Github: Habitat Quality/Threats

Threat data for roads and trials assumed constant
#-------------------------------------------------------------#
#Threat Access Layer
GAP DATA for accessibility of threats: https://www.sciencebase.gov/catalog/item/60259839d34eb12031138e1e

Data:TV.GAP.shp
#-------------------------------------------------------------#
#NLCD Data (buffer)

Data: /Volumes/Research/JodiBrandt/Common/SarahHalperin/Data/HabitatQuality_Data/NLCD
Github: Land use/land use for habitat model 

see NLCDterra.R for creation of data inputs or NLCD_github
#-------------------------------------------------------------#
#General Notes
NOTE: Half saturation constant: based on first run of the model. Should be half of the highest degradation value. 

NOTE: since only one development class I averaged across low, medium, high in the threats and sensitivity layer. All threats/acess remain constant except LULC (i.e. roads, trails, and access)





#------------------------------------------------------------------#
#Compile data for Habitat Model 

```{r}
#Load in Packages
library(rgdal)
library(sp)
library(raster)
library(sf)
library(rgeos)
library(terra)
```

```{r}
#Study area
StudyArea.sf<-st_read("/Volumes/JodiBrandt/Common/SarahHalperin/Data/StudyArea/StudyArea_wgs84.shp")
StudyArea.sf%>%
  st_transform(4087) 

StudyArea.Vec<-terra::vect(StudyArea.sf)

StudyArea.vec.buffer<-buffer(StudyArea.Vec, width=10000)

writeVector(StudyArea.vec.buffer, "/Volumes/JodiBrandt/Common/SarahHalperin/Data/StudyArea/StudyArea_buffer.shp")

#Roads
Idaho.roads.sf<-st_read("/Volumes/Research/JodiBrandt/Common/SarahHalperin/Data/Roads/Idaho_2016_roads/tl_2016_16_prisecroads.shp")

Idaho.roads.sf%>%
  st_transform(4807)

Idaho.roads.vec<-terra::vect(Idaho.roads.sf)

#crop to study area 
TV.roads<-terra::crop(Idaho.roads.vec, StudyArea.vec.buffer)


#Subset by type of road 
TV.primaryRoads<-TV.roads%>%
  subset(TV.roads$MTFCC == "S1100")
TV.secondaryRoads<-TV.roads%>%
  subset(TV.roads$MTFCC == "S1200")

#----------------------------------
#Turn into raster with 0 and 1 values 
#-----------------------------------
StudyArea<-st_as_sf(StudyArea.vec.buffer)

e<-extent(StudyArea)#create blank raster to then use to turn recreation polygon to raster y
r<-raster(nrow=4029, ncol=3343) #taken from raster of a buffer area 
r<-setExtent(r,e)


TV.primaryRoads$rvalue<-1
TV.primaryRoads.sf<-st_as_sf(TV.primaryRoads)

#for some reason cannot get to work with terra 

#primary roads 
TV.pri.raster<-raster::rasterize(TV.primaryRoads.sf, r, field="rvalue") #creates a raster with the extent of the Treasure Valley using NLCD data cropped to TV


TV.secondaryRoads$rvalue<-1
TV.secondaryRoads.sf<-st_as_sf(TV.secondaryRoads)
#secondary roads 
TV.sec.raster<-raster::rasterize(TV.secondaryRoads.sf, r, field="rvalue") #creates a raster with the extent of the Treasure Valley using NLCD data cropped to TV


#Write Files
writeRaster(TV.sec.raster, file="/Volumes/JodiBrandt/Common/SarahHalperin/Data/HabitatQuality_Data/Threats_20012016/sec_roads_c.tif", datatype='INT4S', overwrite=TRUE)

writeRaster(TV.pri.raster, file="/Volumes/JodiBrandt/Common/SarahHalperin/Data/HabitatQuality_Data/Threats_20012016/pri_roads_c.tif", datatype='INT4S', overwrite=TRUE)

#-------------------------------------
#Trails 
#--------------------------------------
#See recreation notes for data sources for trails 
Trail_folder<-"/Volumes/JodiBrandt/Common/SarahHalperin/Data/Recreation/IdahoPR_Trails/"
#no ATV seasonal trails, motorcyclce, motorcycle seasonal, within Treasure Valley 

Trail_folder<-"/Volumes/JodiBrandt/Common/SarahHalperin/Data/Recreation/IdahoPR_Trails/" 
#no ATV seasonal trails, motorcyclce, motorcycle seasonal, within Treasure Valley 

Trail_folder_List<-list.files(Trail_folder, pattern=".shp$", full.names = TRUE) #list files within trails folder

Trails<-lapply(Trail_folder_List,function(i){
  readOGR(i) #bring in all files in input folder
})

TV.trails<-Trails[c(1,3,4,5,6,9)]

#add in urban trails 
urban.sp<-readOGR("/Volumes/JodiBrandt/Common/SarahHalperin/Data/Recreation/Urban_and_Foothill_Ridge_to_River_Trails/Trails.shp")


TV.trails[[1]]$rvalue<-1
TV.trails[[2]]$rvalue<-1
TV.trails[[3]]$rvalue<-1
TV.trails[[4]]$rvalue<-1
TV.trails[[5]]$rvalue<-1
TV.trails[[6]]$rvalue<-1


ID.Trails.r<-lapply(TV.trails, function(i){
  raster::rasterize(i, r, field="rvalue")
})

ID.Trails<-merge(ID.Trails.r[[1]], ID.Trails.r[[2]], ID.Trails.r[[3]], ID.Trails.r[[4]], ID.Trails.r[[5]], ID.Trails.r[[6]])


urban.sp<-readOGR("/Volumes/JodiBrandt/Common/SarahHalperin/Data/Recreation/Urban_and_Foothill_Ridge_to_River_Trails/Trails.shp")
urban.sp$rvalue<-1
urban.r<-raster::rasterize(urban.sp, r, field="rvalue")

ID.Trails.total<-merge(ID.Trails, urban.r)

#Write file
writeRaster(ID.Trails.total,"/Volumes/JodiBrandt/Common/SarahHalperin/Data/HabitatQuality_Data/Threats/Threats_2001/trails_c.tif", datatype='INT4S', overwrite=TRUE)

#-----------------------------------------------_#
#Threat Access Layer
#-----------------------------------------------#

Gap<-st_read("/Volumes/JodiBrandt/Common/SarahHalperin/Data/HabitatQuality_Data/GAP/PADUS2_1Designation_StateID.shp")

 #assign access based on status code 
Gap$ID<-seq.int(nrow(Gap))
Gap.select<-Gap[, c(30,43)]
Gap.select$GAP_Sts[Gap.select$GAP_Sts == 1]<-0
Gap.select$GAP_Sts[Gap.select$GAP_Sts == 2]<-.333333
Gap.select$GAP_Sts[Gap.select$GAP_Sts == 3]<-.666666
Gap.select$GAP_Sts[Gap.select$GAP_Sts == 4]<-1
Gap.select<-st_make_valid(Gap.select) #make sure valid polygon
Gap.select$ACCESS<-Gap.select$GAP_Sts

Gap.select.vec<-vect(Gap.select) #turn to spacvector

TV.Gap.vec<-terra::crop(Gap.select.vec, StudyArea.vec.buffer) #crop 

#write file
writeVector(TV.Gap.vec, "/Volumes/JodiBrandt/Common/SarahHalperin/Data/HabitatQuality_Data/GAP/TV.GAP.shp")


```

