---
title: "Chapter 1 Spearman's Correlation"
author: "Sarah Halperin"
date: "7/6/2021"
output: pdf_document
---
```{r}
  library(RColorBrewer)
  library(ggplot2)
  library(raster)
  library(rgdal)
  library(sf)
  library(dplyr)
  library("FactoMineR")
  library(factoextra)
  library(reshape2)
  library(tidyverse)
  library(fmsb)
  library(Hmisc)
  library(corrplot)
  library(RColorBrewer)

```

##Overview 

This RMD runs a spearman's correlation analysis and PCA on the ES layers and PVR. 


#Bring in data 
```{r echo=FALSE, message=FALSE}

Folder<-"/Volumes/JodiBrandt/Common/SarahHalperin/"


ES_dataframe<-read.csv(paste0(Folder, "Data/ES_Spatial_Layers/ESDataFrame2016_Full.csv"), header=T)
#ES_dataframe<-read.csv(paste0(Folder, "Data/ES_CSV/02_ESDataFrame2016_Full.csv"), header=T)
ES_dataframe$X<-NULL #I think might be because how exporting. 


#ES_dataframe$LandUseW<-ifelse(is.na(ES_dataframe$Carbon), NA ,ES_dataframe$LandUseW) #make sure NA for study region
ES_dataframe$PVR.Group<-ifelse(is.na(ES_dataframe$PVR.Group),"None", ES_dataframe$PVR.Group) #assign None to NA values 
ES_dataframe<-na.omit(ES_dataframe) #need to have no NA for PCA. For ES that weren't removed in Set_up script.
ES_dataframe$LandUse<-NULL


set.seed(123)
ES_dataframe.scale<-scale(ES_dataframe[,1:6]) #want default settings 
ES_dataframe.scale<-as.data.frame(ES_dataframe.scale, as.numeric)

#scale so mean is close to zero and std is 1. important for PCA and cluster 
```

#Run Spearman's correlation analysis (Karimi et al. 2021)
#http://bwlewis.github.io/covar/missing.html

```{r}
#http://www.sthda.com/english/wiki/correlation-matrix-a-quick-start-guide-to-analyze-format-and-visualize-a-correlation-matrix-using-r-software

ES_dataframe<-ES_dataframe[1:7]
ES_dataframe<-ES_dataframe[,c("FoodProvision", "LivestockCropProduction", "Carbon", "NitrogenRetention", "Habitat", "Recreation", "PVR")]

colnames(ES_dataframe)<-c("Food Provision", "Livestock Crop Production", "Carbon Sequestration", "Nitrogen Retention", "Habitat Quality", "Recreation", "PVR")

spearman<-cor(ES_dataframe[,1:6], method="spearman", use="complete") #run the spearman's correlation. 

spearman.pvr<-cor(ES_dataframe[,1:7], method="spearman", use="complete") #check which column for PVR 


#correlation plot 
cor.sig<-rcorr(as.matrix(ES_dataframe.scale), type="spearman")
cor.sig$P



corrplot(spearman.pvr, 
         method="number", 
         type = "upper", 
        col = colorRampPalette(c("white", "deepskyblue", "blue4"))(100))


#minimum maximimum normalization 
norm_minmax<-function(x){
  (x-min(x))/(max(x)-min(x))
}

#Plot XY ES and PVR visualization plots 
df.cor.s<-ES_dataframe
df.cor.s <- df.cor.s %>%
  mutate_at(vars(1:6), norm_minmax)


colnames(df.cor.s)<-c("Food Provision", "Livestock Crop Production", "Carbon Sequestration","Nitrogen Retention" , "Habitat", "Recreation", "PVR")

df.cor.m<-melt(df.cor.s, id="PVR") 

set.seed(123)
df.cor.m<-slice_sample(df.cor.m, n=10000)


df.cor.m$variable<-ordered(df.cor.m$variable, levels=c("Food Provision", "Livestock Crop Production", "Carbon Sequestration","Nitrogen Retention" , "Habitat", "Recreation"))


ggplot(df.cor.m, aes(x=PVR, y=value))+
  geom_point()+
  facet_grid(. ~ variable)+
  ylab("Standardized ES Supply")+
  xlab("PVR value")
```

#PCA Analysis 
http://www.sthda.com/english/articles/31-principal-component-methods-in-r-practical-guide/118-principal-component-analysis-in-r-prcomp-vs-princomp/#compute-pca-in-r-using-prcomp

https://bioconductor.org/packages/release/bioc/vignettes/PCAtools/inst/doc/PCAtools.html
karimi et al. 2021 spearman rank because non-normal distribution. Cor test for significant factor analysis with varimax rotation. PCA chose eigenvalues that were greater than 1. NEED TO CHECK THAT IS WHAT I AM DOING. 

https://www.datacamp.com/community/tutorials/pca-analysis-r
http://www.sthda.com/english/wiki/factoextra-r-package-easy-multivariate-data-analyses-and-elegant-visualization


```{r}
#PCA
#https://www.datacamp.com/community/tutorials/pca-analysis-r

ES_dataframe <- ES_dataframe %>%
   mutate ( PVR.Group = fct_relevel(PVR.Group, "high", "med", "low" )) %>%
  mutate( PVR.Group = recode(PVR.Group, high='High', med='Medium', low='Low')) 

colnames(ES_dataframe)<-c("Food Provision", "Livestock Crop Production","Carbon", "Recreation",  "Nitrogen Retention", "Habitat", "PVR","LandUse", "PVR.Group", "LandUseW")
  
set.seed(123)
ES_dataframe.s<-sample_n(ES_dataframe, 10000)


res.pca<-PCA(ES_dataframe.s[1:6], scale.unit=TRUE, graph=FALSE)

PVR_color<-c("#165d43","#788f33","#aeac4c")


P<-factoextra::fviz_pca_biplot(res.pca, 
                            repel = TRUE,
                            label = "var",
                            col.var = "black",
                            palette= PVR_color,
                            col.ind = ES_dataframe.s$PVR.Group, 
                            habillage = as.factor(ES_dataframe.s$PVR.Group), 
                            pointsize = 3,
                            addEllipses = TRUE,
                            ellipse.level=0.95, 
                            legend.title="Agricultural Land Quality"
                            )+
  scale_shape_manual(values=c(1,0,2))

pdf(paste0(Folder, "PCRBiplot.pdf"),         # File name
    width = 8, height = 7, # Width and height in inches
    bg = "white",          # Background color
    colormodel = "cmyk",    # Color model (cmyk is required for most publications)
    paper = "A4")          # Paper size

P

dev.off()
    
summary(res.pca) 
get_eig(res.pca)
#-------------------------------------#

#-------------------------------------#
#Plots
factoextra::fviz_eig(res.pca) #scee plot which shows the percentage of variances explained by each principal component

factoextra::fviz_screeplot(res.pca, addlabels=TRUE) #alternative scree plot 


#extract the results for variables 
var<-factoextra::get_pca_var(res.pca)
var$contrib #contriution of variables 

#factoextra::fviz_pca_ind(res.pca, #individual 
                       #)

factoextra::fviz_pca_var(res.pca,
             )

#contributions of variables to PC1
factoextra::fviz_contrib(res.pca, choice = "var", axes= 1) 
#contribution of variables to PC2
factoextra::fviz_contrib(res.pca, choice = "var", axes= 2) 
#contribution of variables to PC3
factoextra::fviz_contrib(res.pca, choice = "var", axes= 3) 
#-------------------------------------#

```


