---
title: "Chapter 1 Set up"
author: "Sarah Halperin"
date: "6/29/21"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

##Overview 

This RMD file creates a dataframe from the ES layers for use in the analyses. The script repeats for years 2016, 2030, and 2050. 

```{r echo=FALSE, message=FALSE}
library(raster) # read and edit rasters
library(Hmisc)  # useful functions for data analysis
library(rgdal)  # Add this package to read .tif files
library(dplyr) 
library(reshape)
library(sf)
library(terra)
```


#Generate Data Frame For 2016

##Bring in baseline data 
```{r}
Folder<-"/Volumes/JodiBrandt/Common/SarahHalperin/" 
StudyArea.vec<-terra::vect("/Volumes/JodiBrandt/Common/SarahHalperin/Data/StudyArea/StudyArea_wgs84.shp")

```

## Bring in 2016 data 
```{r} 
NLCD_2016<-terra::rast("/Volumes/JodiBrandt/Common/SarahHalperin/Data/LandUse/TV_NLCD/NLCD_Weighted/TV.NLCD_2016w.tif")

List_2016 <-list(
  Food=terra::rast(paste0(Folder,"Data/Cropscape_Data/SpatialYieldLayers/Food_2016.tif")),
  Pasture=terra::rast(paste0(Folder, "Data/Cropscape_Data/SpatialYieldLayers/Pasture_2016.tif")),
  Carbon=terra::rast(paste0(Folder,"InVEST/Carbon/09-30-21Output/20012016/tot_c_fut_weight.tif")), #Year 2016 
  Recreation=terra::rast(paste0(Folder, "Data/Recreation/RecreationProvision/02_recreation_2016sum.tif")),
  Retention=terra::rast("/Volumes/JodiBrandt/Common/SarahHalperin/InVEST/WaterQuality/NLCD2016w/Retention2016_chaplin30m.tif"),
  Habitat=terra::rast("/Volumes/JodiBrandt/Common/SarahHalperin/InVEST/HabitatQuality/20012016/quality_f_02.tif"), 
  PVR=terra::rast(paste0(Folder, "Data/AFT_Data/PVR_TV.tif"))
)

#Reproject 
List_2016.p<-lapply(List_2016, FUN = function(x) {
  terra::project(x, NLCD_2016)
})

StudyArea.vec<-terra::project(StudyArea.vec, NLCD_2016)

#Crop 
List_2016.p<-lapply(List_2016.p, FUN = function(x) {
  terra::crop(x, StudyArea.vec)
  return(terra::mask(x, StudyArea.vec))
})
```


##Create Dataframe
```{r}
Stack<-c(List_2016.p$Food, List_2016.p$Pasture, List_2016.p$Carbon, List_2016.p$Recreation, List_2016.p$Retention, List_2016.p$Habitat, List_2016.p$PVR, NLCD_2016)

Brick<-brick(Stack) #sometimes need to run a couple of times to get it to work 

ES_dataframe<-as.data.frame(Brick)


names(ES_dataframe)<-c("FoodProvision","LivestockCropProduction","Carbon","Recreation", "NitrogenRetention", "Habitat",  "PVR", "LandUse")

#remove NA if carbon is NA, assumed carbon is entire study region as it is an inVEST output  
ES_dataframe<-ES_dataframe %>% mutate_all(~ifelse(is.nan(.), NA, .))
ES_dataframe<-ES_dataframe[!is.na(ES_dataframe$Carbon),]



#replace NA values for Food, Livestock
ES_dataframe$FoodProvision<-ifelse(is.na(ES_dataframe$FoodProvision), 0, ES_dataframe$FoodProvision)
ES_dataframe$LivestockCropProduction<-ifelse(is.na(ES_dataframe$LivestockCropProduction), 0, ES_dataframe$LivestockCropProduction)
```

#Join ES values with Land use categories and PVR values. Group PVR values into low, med, high 
```{r}
#TO INCLUDE LAND USE
LandUseW<-c("Open Water","Cultivated Crops", "Development","Barren land", "Pasture/Hay", "Shrub", "Grassland", "Mergent Herbaceous", "Woody Wetland", "Evergreen Forest", "Deciduous Forest", "Mixed Forest")
LandUse<-c(11,82,25,31, 81,52, 71, 95, 90, 42,41,43)
LandUseCategories<-as.data.frame(cbind(LandUseW, LandUse))
LandUseCategories$LandUse<-as.numeric(LandUseCategories$LandUse)




#Determine 1/3 quantiles 
PVR.test<-as.data.frame(List_2016.p$PVR) #run in 2016 data 
quantile(PVR.test$PVR, probs=seq(0,1,1/3), na.rm=TRUE) #0.22, 0.49
median(PVR.test$PVR, na.rm=TRUE) #0.36

labels=c('low','med','high')

ES_dataframe<-ES_dataframe %>%
  mutate(PVR.Group = 
           cut(PVR, breaks = c(0,0.22,0.49,1), right = F, labels=labels))


ES_dataframe.F<-full_join(ES_dataframe, LandUseCategories, by="LandUse")



write.csv(ES_dataframe.F,paste0(Folder, "Data/ES_CSV/02_ESDataFrame2016_Full.csv"), row.names=FALSE)
```

#----------------------------------------------------------------------------------------------------#
##Bring in 2030 Data
```{r}
#2030
NLCD_2030<-terra::rast("/Volumes/JodiBrandt/Common/SarahHalperin/Data/LandUse/TV_NLCD/NLCD_Weighted/TV.NLCD_2030.tif")

List_2030 <- list(
  Food=terra::rast(paste0(Folder,"Data/Cropscape_Data/SpatialYieldLayers/Food_2030.tif")), 
  Pasture=terra::rast(paste0(Folder, "Data/Cropscape_Data/SpatialYieldLayers/Pasture_2030.tif")),
  Carbon = terra::rast(paste0(Folder,"InVEST/Carbon/09-30-21Output/20302050/tot_c_cur_weight.tif")), 
  Recreation = terra::rast(paste0(Folder, "Data/Recreation/RecreationProvision/02_recreation_2030sum.tif")),
  Retention= terra::rast("/Volumes/JodiBrandt/Common/SarahHalperin/InVEST/WaterQuality/NLCD2030/Retention2030_chaplin30m.tif"),
  Habitat= terra::rast("/Volumes/JodiBrandt/Common/SarahHalperin/InVEST/HabitatQuality/20302050/quality_c_02.tif"),
  PVR= terra::rast(paste0(Folder, "Data/AFT_Data/PVR2030.tif"))
) 


#Reproject 
List_2030.p<-lapply(List_2030, FUN = function(x) {
  terra::project(x, NLCD_2030)
})

StudyArea.vec<-terra::project(StudyArea.vec, NLCD_2030)

#Crop 
List_2030.p<-lapply(List_2030.p, FUN = function(x) {
  terra::crop(x, StudyArea.vec)
  return(terra::mask(x, StudyArea.vec))
})
```


##Create Dataframe 2030
```{r}
Stack<-c(List_2030.p$Food, List_2030.p$Pasture, List_2030.p$Carbon, List_2030.p$Recreation, List_2030.p$Retention, List_2030.p$Habitat, List_2030.p$PVR, NLCD_2030)

Brick<-brick(Stack) #sometimes need to run a couple of times to get it to work 

ES_dataframe<-as.data.frame(Brick)


names(ES_dataframe)<-c("FoodProvision","LivestockCropProduction","Carbon","Recreation", "NitrogenRetention", "Habitat",  "PVR", "LandUse")

#remove NA if carbon is NA, assumed carbon is entire study region as it is an inVEST output  
ES_dataframe<-ES_dataframe %>% mutate_all(~ifelse(is.nan(.), NA, .))
ES_dataframe<-ES_dataframe[!is.na(ES_dataframe$Carbon),]



#replace NA values for Food, Livestock
ES_dataframe$FoodProvision<-ifelse(is.na(ES_dataframe$FoodProvision), 0, ES_dataframe$FoodProvision)
ES_dataframe$LivestockCropProduction<-ifelse(is.na(ES_dataframe$LivestockCropProduction), 0, ES_dataframe$LivestockCropProduction)
```

#Join ES values with Land use categories and PVR values. Group PVR values into low, med, high 
```{r}
#TO INCLUDE LAND USE
LandUseW<-c("Open Water","Cultivated Crops", "Development","Barren land", "Pasture/Hay", "Shrub", "Grassland", "Mergent Herbaceous", "Woody Wetland", "Evergreen Forest", "Deciduous Forest", "Mixed Forest")
LandUse<-c(11,82,25,31, 81,52, 71, 95, 90, 42,41,43)
LandUseCategories<-as.data.frame(cbind(LandUseW, LandUse))
LandUseCategories$LandUse<-as.numeric(LandUseCategories$LandUse)

labels=c('low','med','high')

ES_dataframe<-ES_dataframe %>%
  mutate(PVR.Group = 
           cut(PVR, breaks = c(0,0.22,0.49,1), right = F, labels=labels))


ES_dataframe.F<-full_join(ES_dataframe, LandUseCategories, by="LandUse")



write.csv(ES_dataframe.F,paste0(Folder, "Data/ES_CSV/02_ESDataFrame2030_Full.csv"), row.names=FALSE)
```


##Bring in 2050 Data
```{r}
#2050
NLCD_2050<-terra::rast("/Volumes/JodiBrandt/Common/SarahHalperin/Data/LandUse/TV_NLCD/NLCD_Weighted/TV.NLCD_2050.tif")

List_2050 <- list( 
  Food=terra::rast(paste0(Folder,"Data/Cropscape_Data/SpatialYieldLayers/Food_2050.tif")),
Pasture=terra::rast(paste0(Folder, "Data/Cropscape_Data/SpatialYieldLayers/Pasture_2050.tif")),
Carbon=terra::rast(paste0(Folder,"InVEST/Carbon/09-30-21Output/20302050/tot_c_fut_weight.tif")), 
Recreation=terra::rast(paste0(Folder, "Data/Recreation/RecreationProvision/02_recreation_2050sum.tif")),
Retention=terra::rast("/Volumes/JodiBrandt/Common/SarahHalperin/InVEST/WaterQuality/NLCD2050/Retention2050_chaplin30m.tif"),
Habitat=terra::rast("/Volumes/JodiBrandt/Common/SarahHalperin/InVEST/HabitatQuality/20302050/quality_f_02.tif"),
PVR=terra::rast(paste0(Folder, "Data/AFT_Data/PVR2050.tif"))
)

#Reproject 
List_2050.p<-lapply(List_2050, FUN = function(x) {
  terra::project(x, NLCD_2050)
})

StudyArea.vec<-terra::project(StudyArea.vec, NLCD_2050)

#Crop 
List_2050.p<-lapply(List_2050.p, FUN = function(x) {
  terra::crop(x, StudyArea.vec)
  return(terra::mask(x, StudyArea.vec))
})

```


##Create Dataframe 2050
```{r}
Stack<-c(List_2050.p$Food, List_2050.p$Pasture, List_2050.p$Carbon, List_2050.p$Recreation, List_2050.p$Retention, List_2050.p$Habitat, List_2050.p$PVR, NLCD_2050)

Brick<-brick(Stack) #sometimes need to run a couple of times to get it to work 

ES_dataframe<-as.data.frame(Brick)


names(ES_dataframe)<-c("FoodProvision","LivestockCropProduction","Carbon","Recreation", "NitrogenRetention", "Habitat",  "PVR", "LandUse")

#remove NA if carbon is NA, assumed carbon is entire study region as it is an inVEST output  
ES_dataframe<-ES_dataframe %>% mutate_all(~ifelse(is.nan(.), NA, .))
ES_dataframe<-ES_dataframe[!is.na(ES_dataframe$Carbon),]



#replace NA values for Food, Livestock
ES_dataframe$FoodProvision<-ifelse(is.na(ES_dataframe$FoodProvision), 0, ES_dataframe$FoodProvision)
ES_dataframe$LivestockCropProduction<-ifelse(is.na(ES_dataframe$LivestockCropProduction), 0, ES_dataframe$LivestockCropProduction)
```

#Join ES values with Land use categories and PVR values. Group PVR values into low, med, high 
```{r}
#TO INCLUDE LAND USE
LandUseW<-c("Open Water","Cultivated Crops", "Development","Barren land", "Pasture/Hay", "Shrub", "Grassland", "Mergent Herbaceous", "Woody Wetland", "Evergreen Forest", "Deciduous Forest", "Mixed Forest")
LandUse<-c(11,82,25,31, 81,52, 71, 95, 90, 42,41,43)
LandUseCategories<-as.data.frame(cbind(LandUseW, LandUse))
LandUseCategories$LandUse<-as.numeric(LandUseCategories$LandUse)

labels=c('low','med','high')

ES_dataframe<-ES_dataframe %>%
  mutate(PVR.Group = 
           cut(PVR, breaks = c(0,0.22,0.49,1), right = F, labels=labels))


ES_dataframe.F<-full_join(ES_dataframe, LandUseCategories, by="LandUse")



write.csv(ES_dataframe.F,paste0(Folder, "Data/ES_CSV/02_ESDataFrame2050_Full.csv"), row.names=FALSE)
```

##Export raster layers to look at them in Arc or Qgis 
```{r}
#2016
ES_folder<-paste0(Folder,"Data/ES_Spatial_Layers/")

ES_names.2016<-c("Food_2016", "Pasture_2016", "Carbon_2016", "Recreation_2016", "Retention_2016", "Habitat_2016")

mapply(function(x,n){
  writeRaster(x, file=paste0(ES_folder,n,".tif"), overwrite=TRUE)
}, List_2016.p, ES_names.2016)

#2030 
ES_names.2030<-c("Food_2030", "Pasture_2030", "Carbon_2030", "Recreation_2030", "Retention_2030", "Habitat_2030")

mapply(function(x,n){
  writeRaster(x, file=paste0(ES_folder,n,".tif"), overwrite=TRUE)
}, List_2030.p, ES_names.2030)

#2050
ES_names.2050<-c("Food_2050", "Pasture_2050", "Carbon_2050", "Recreation_2050", "Retention_2050", "Habitat_2050")

mapply(function(x,n){
  writeRaster(x, file=paste0(ES_folder,n,".tif"), overwrite=TRUE)
}, List_2050.p, ES_names.2050)
```

#------------------------------------------------------------------------------------#
## NOTE: Assign 0 values to NA values for Food, Pasture. As these aren't values we don't have but areas that do not have this ES. 


